---
title: "MTT#1 Cell Density"
author: "Andrés Gordo"
date: "2023-06-23"
output:
  pdf_document: 
    fig_caption: yes
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Tidying the data

First, I load the libraries needed and read the raw data from **Omega** in *.xlsx*. Then I create my **tidy data** selecting the proper values and assigning them to their corresponding **cell density**.

```{r,message=FALSE, warning=FALSE, echo=TRUE}
library(ggpubr)
library(readxl)
library(caret)
mtt_raw <- read_excel("C:\\Users\\andre\\OneDrive - unizar.es\\Laboratorio\\UCL ILDH 2023\\Results\\MTT\\MTT1_ANDRES_PDX_236.xlsx", col_names = FALSE)
mtt_tidy<-data.frame(absorbance=unlist(mtt_raw[4:9,3:10]), density=0)
mtt_tidy<-mtt_tidy[1:45,] #This assigns the absorbances to the cell density
mtt_tidy[1:12, 2]<- 2000
mtt_tidy[13:24, 2]<- 5000
mtt_tidy[25:36, 2]<- 8000
mtt_tidy[37:45, 2]<- 10000
```

## Plotting with ggplot

Violin plot with boxplot, adding a two sample *t test* for significance.

```{r, echo=TRUE}

density_comparisons<-list( c("2000", "5000"), 
                         c("5000", "8000"),c("8000", "10000"))
model<-lm(density~absorbance, mtt_tidy[1:36,])
violin<- ggviolin(mtt_tidy,
          x          = "density", 
          y          = "absorbance", 
          title      = "MTT#1 Results & Analysis",
          xlab       = "Number of Cells seeded",
          ylab       = "Absorbance",
          fill       = "density",
          palette = c("#c994c7", "#df65b0", "#980043", "#67001f"),
          alpha      = 0.8,
          add        = "boxplot",
          add.params = list(fill = "white")) +
  stat_compare_means(comparisons = density_comparisons, 
                     label       = "p.signif", method="t.test")
 
  
violin
```

## Density checking

While preparing the dilutions for the MTT I took less RPMI volume than I should have for the 10k cell density solution. Due to that, I could not fill all the wells, only having 9 compared to the 12 wells of the rest of the groups (2k, 5k, 8k). That means that there should be more than 10k cell per well. If I create a regression model trained with the 2k, 5k and 8k only, and then predict the cell density corresponding to the absorbance shown in the 10k well I should be able to ched whether the hypothesis is true.

```{r, echo=TRUE, include=TRUE}
model<-lm(density~absorbance, mtt_tidy[1:36,]) #I create the model
prediction<-predict(model, mtt_tidy[37:45,]) #I predict the density 
print(paste(c(round(mean(prediction)),"> 10k"), collapse = " ")) 
#I print the mean of those predicted values

```

Indeed, *I seeded more cells per well in the 10k group than I should have*.
